Databrary: A digital data library for sharing research video

### 2015-01-18 jcdl-15/paper.tex
some more notes, clean out stuff that has been addressed, some edits to paper and references

Databrary can incorporate heterogeneous datasets within our target domain by allowing secondary, general-purpose flat-file data, materials, and references to be associated with the primary data.
This facilitates a depth of access to configure data more to the original context in which it was collected and analyzed.
% that use of "configure" doesn't make sense to me -dylan
At the same time, it provides a common, standard structure for recordings and their associated metadata, including context, contents, and annotations.
We achieve this conformance of datasets in the library through a combination of in-depth investigative curation and ingest processes for already-collected data, and a well-defined user experience that provides a platform for researchers to contribute and organize all of their research data as they are collected.
The following sections provide specific detail on the design and decision-making behind this unifying structure.
%%dg more like a system informed by the pitfalls of committing solely to either extreme.

@@

%%dg would a figure illustration be helpful here? do we already have one?

@@

% should we actually use the term "record"?  Alternatives: "grouping", "label", "annotation", ...

@@

Researchers record activities with individuals either for the purpose of a later, closer review or analysis, or simply to document and verify measurements and observations made during the collection.
Thus, the recording itself is often the core item of research, representing a single data collection, experiment, trial, visit, or case depending on the nature of the research.
However, in many cases a single session can comprise multiple videos: from serial data streams due to recording equipment being producing multiple files, and/or in parallel from multiple data capture sources operating at the same time.
These primary recordings may also be augmented with other forms of timeseries data such as motion capture, eye tracking, or physiological recordings, as well as with additional point-in-time measurements, observations, or surveys, all of which have specific and critical temporal relationships with the primary data.

% should we be using more language like, "we observed", "in practice"?
When dealing with multiple primary recordings, researchers often reduce these raw files into a single representation.
For example, they may concatenate serial recordings or make a composite video by joining or overlaying frames.
Secondary data is also often overlaid or combined in this way, such as by adding an additional waveform or frequency visualization below a video, or by overlaying crosshairs representing location or gaze data.
Doing so allows for more convenient human analysis of the data in standard tools that only support a single data stream, but also loses some of the raw data, by obscuring, masking, or downsampling spatially or temporally.

@@

%%dg, i guess it's crucial to make sure the audience understands that one of the specific challenges for us is that we're not just storing comprehensive single video files, but multiple files that capture the same phenomenon and how do we keep all those together in a meaningful way...to differentiate us from being "just another you tube" .. which i guess is what is already being said above. Though might need to do a little more hand holding for the reader as to why this is an issue in general and why it's important to even those who might be faced with different types of video data to collect (? maybe I'm rambling now ?).

@@

An object called a record serves a number of purposes around session and metadata organization.
A record itself represents any number of variables or measurements---what would traditionally be a row of a database or flat file.
For example, a record may represent an individual participant within the dataset and contain their subject identifier, gender, birthdate, and any other demographic, survey, or measurement data about that participant.
A record could also represent a particular experimental task, condition, location, or outcome, and contain variables with corresponding parameters or descriptions.
Records are thus labeled and divided by the type of entity they represent.
The variables themselves are textual, numeric, date, or potentially any other scalar data as in any traditional tabular data interface, with the records of a given type constituting the tabular rows.

@@

%%dg other notes: should we give a bit more discussion of the infrastruction - Postgresql, ffmpeg, using HPC for transcoding, storage, frontend technologies etc. 
% i'm not sure. do other papers have this sort of "methods" section? -dylan

@@

As the extant research data libraries demonstrate, there is always the tension between standardization and flexibility.
Since we're primarily dealing with video and audio data, we have, without much loss of flexibility, standardized on a particular encoding format (codecs and parameters) that allows for convenient access and HTML5-based streaming, while still allowing arbitrary resolutions and sample/frame-rates.
We also allow for ``dumps'' using certain portable data formats to include any data that cannot be encompassed by our standard formats, reasoning that these data may be relevant for re-use, but not broadly adopted by the community.

Here we have also introduced a number of metadata concepts where this tension becomes most apparent, namely the available record types, variables (and values within categorical data types), and tag keywords.
We have adopted an approach of encouraging adherence to existing standardized options but also allowing new types to be created as the data require it.
We encourage standardization both on-line as data are collected and entered via our user interface using auto-completion, suggestions, and drop-down options, and off-line via manual review and curaion of contributed data.

>>>>

Databrary can incorporate heterogeneous datasets within our target domain by allowing secondary, general-purpose flat-file data, materials, and references to be associated with the primary data.
This facilitates storage and retrieval of data more closely mapped to the original context in which it was collected and analyzed.
At the same time, it provides a common, standard structure for recordings and their associated metadata, including context, contents, and annotations.
We achieve this conformance of datasets in the library through a combination of in-depth investigative curation and ingest processes for already-collected data, and a well-defined user experience that provides a platform for researchers to contribute and organize all of their research data as they are collected.
The following sections provide specific detail on the design and decision-making behind this unifying structure.
%%dg more like a system informed by the pitfalls of committing solely to either extreme.

@@

%%dg would a figure illustration be helpful here? do we already have one?
%%dg in general, how much room do we have for screenshots, if they help explain?

@@

% should we actually use the term "record"?  Alternatives: "grouping", "label", "annotation", ...
% as long as we're clear in defining it. the record section appears to be doing this.

@@

%dg "The session was designed both to organize videos in a way that is familiar to researchers, as well as to allow them to 'go into' a session and add or edit data." - stuff like this to remind the reader why we're discussing these components in specific (rather than some other aspect of the system).
Researchers record activities with individuals either for the purpose of a later, closer review or analysis, or simply to document and verify measurements and observations made during the collection.
Thus, the recording itself is often the core item of research, representing a single data collection, experiment, trial, visit, or case depending on the nature of the research.
However, in many cases a single session can comprise multiple videos: from serial data streams due to recording equipment being producing multiple files, and/or in parallel from multiple data capture sources operating at the same time.
These primary recordings may also be augmented with other forms of timeseries data such as motion capture, eye tracking, or physiological recordings, as well as with additional point-in-time measurements, observations, or surveys, all of which have specific and critical temporal relationships with the primary data.

% should we be using more language like, "we observed", "in practice"?
%%dg - if it helps to illustrate particular points better, sure. 
When working with multiple primary recordings, researchers often reduce these raw files into a single representation.
For example, they may concatenate serial recordings or make a composite video by joining or overlaying frames.
Secondary data is also often overlaid or combined in this way, such as by adding an additional waveform or frequency visualization below a video, or by overlaying crosshairs representing location or gaze data.
Doing so allows for more convenient human analysis of the data in standard tools that only support a single data stream, but also loses some of the raw data, by obscuring, masking, or downsampling spatially or temporally.

@@

%%dg, i guess it's crucial to make the audience aware that one of the specific challenges for us is that we're not just storing comprehensive single video files, but multiple files that capture the same phenomenon and how do we keep all those together in a meaningful way...to differentiate us from being "just another you tube".

@@

%%dg maybe add some langauge to the beginning of this section that is explicit that the "record" is designed with the flexibility we talk about above in mind.
An object called a record serves a number of purposes around session and metadata organization. 
A record itself represents any number of variables or measurements---what would traditionally be a row of a database or flat file.
For example, a record may represent an individual participant within the dataset and contain their subject identifier, gender, birthdate, and any other demographic, survey, or measurement data about that participant.
A record could also represent a particular experimental task, condition, location, or outcome, and contain variables with corresponding parameters or descriptions.
Records are thus labeled and divided by the type of entity they represent.
The variables themselves are textual, numeric, date, or potentially any other scalar data as in any traditional tabular data interface, with the records of a given type constituting the tabular rows.

@@

%%dg other notes: should we give a bit more discussion of the infrastruction - Postgresql, ffmpeg, using HPC for transcoding, storage, frontend technologies etc. 
% i'm not sure. do other papers have this sort of "methods" section? -dylan
%%dg - other papers in this general area (see: Peer_2012, Griffin_etal_2014, and Witt_2012 from references) do comment on the technology used (if there actually is a product and not just a proposed framework) to varying degrees. I don't think we have to spend more than a paragraph on it, but it would be good to mention the stack in terms of taking, augmenting, and storing video files and metadata overall.

@@

As the extant research data libraries demonstrate, there is always the tension between standardization and flexibility.
Since we're primarily dealing with video and audio data, we have, without much loss of flexibility, standardized on a particular encoding format (codecs and parameters) that allows for convenient access and HTML5-based streaming, while still allowing arbitrary resolutions and sample/frame-rates.
We also allow for ``dumps'', which allow users - using certain portable data formats (e.g. ...) - to include any data that cannot be encompassed by our standard formats. In this way, we can incorporate data which may be relevant for re-use, but not broadly adopted by the community, without making the system too rigid.

Here we have also introduced a number of metadata concepts where this tension becomes most apparent, namely the available record types, variables (and values within categorical data types), and tag keywords.
We have adopted an approach of encouraging adherence to existing standardized options but also allowing new types to be created as the data require it.
We encourage standardization both on-line as data are collected and entered via our user interface using auto-completion, suggestions, and drop-down options, and off-line via manual review and curation of contributed data.
